---
title: "palmpeng"
format: html
editor: visual
---

```{r}
# Load tidyverse and palmerpenguins libraries
library(tidyverse)
library(palmerpenguins)
  

# Create a ggplot scatter plot using the penguins dataset. 
# Plot bill_length_mm on the x-axis and bill_depth_mm on the y-axis.
# Color points by species and use geom_point with size 3.
# Add a minimal theme and clean labels.
plot <- ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_point(size = 3) +
  labs(title = "Bill Length vs Bill Depth of Penguins",
       x = "Bill Length (mm)",
       y = "Bill Depth (mm)") +
  theme_minimal()

plot

```

```{r}

#create boxplot comparing flipper_length_mm across different species, faceted by island, using custom color palettes.
boxplot <- ggplot(data = penguins, aes(x = species, y = flipper_length_mm, fill = species)) +
  geom_boxplot() +
  facet_wrap(~ island) +
  labs(title = "Flipper Length by Species and Island",
       x = "Species",
       y = "Flipper Length (mm)") +
  theme_minimal() +
  scale_fill_manual(values = c("Adelie" = "blue", "Chinstrap" = "red", "Gentoo" = "green"))

boxplot

```

```{r}
#build a scatter plot mapping body_mass_g against flipper_length_mm, adding a linear regression trend line (geom_smooth) for each species.
scatter_plot <- ggplot(data = penguins, aes(x = flipper_length_mm, 
                                            y = body_mass_g, 
                                            color = species)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Body Mass vs Flipper Length of Penguins",
       x = "Flipper Length (mm)",
       y = "Body Mass (g)") +
  theme_minimal()

scatter_plot

```

```{r}
#installing chatrr
    #install.packages("ellmer")
    #install.packages("chattr")

usethis::edit_r_environ()

usethis::edit_r_profile()

# Option B: For OpenAI ChatGPT
options(.chattr_chat = ellmer::chat_openai())

```

```{r}

#Show me how to load the built-in iris dataset and use ggplot2 to build a scatter plot of Sepal.Length versus Sepal.Width colored by Species.
data(iris)
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point() +
  labs(title = "Sepal Length vs Sepal Width",
       x = "Sepal Length",
       y = "Sepal Width") +
  theme_minimal()

```

```{r}
# Simple Example: Checking Sea Surface Temperature (SST)
sst <- 30.2

if (sst > 29.5) {
  print("Warning: Marine heatwave threshold exceeded!")
} else {
  print("SST remains within baseline parameters.")
}

#if else 
# Load tidyverse (if not already loaded)
library(tidyverse)

# Sample coral data
coral_monitoring <- tibble(
  site = c("Site_A", "Site_B", "Site_C"),
  depth_m = c(12, 35, 8)
)

# Classify depth using if_else
coral_monitoring <- coral_monitoring %>% 
  mutate(zone = if_else(depth_m > 30, "Deep Reef", "Shallow Reef"))

# mutate()
#%>%

#case when **tildes

coral_monitoring <- coral_monitoring %>% 
  mutate(reef_category = case_when(
    depth_m < 10  ~ "Lagoon / Flats",
    depth_m <= 30 ~ "Crest / Slope",
    depth_m > 30  ~ "Mesophotic / Deep",
    TRUE          ~ "Unclassified" # Catch-all remainder
 ))

```

Using the chattr app or inline copilot, your task here is to write an R script that initialises a tibble named marine_stations containing a column for salinity with values 35, 28, 32, and 12. Use the pipe operator (%>%) and mutate() combined with case_when() to create a new column named environment_type. Classify values below 15 as "Estuarine", values between 15 and 30 as "Brackish", and values above 30 as "Marine". Run your script and verify the table structure.
```{r}

#Marine_stations as a tibble containing a column for salinity with values 35, 28, 32, and 12.

marine_stations <- tibble(
  salinity = c(35, 28, 32, 12)
)

#Use %>% and mutate(), along with case_when to create a new column named environmental_type.

envrionment_type <- marine_stations %>%
  mutate(environment_type = case_when(
    #classify "Estuarine " below 15, "Brackish" between 15 and 30, and "Marine" above 30.
    salinity < 15 ~ "Estuarine",
    salinity >= 15 & salinity <= 30 ~ "Brackish",
    salinity > 30 ~ "Marine"
  ))


```


```{r}
##BASE R 
# Classic loop anatomy
for (year in 2020:2024) {
  print(paste("Processing climate data for year:", year))
}

# Loop across index sequences
transect_lengths <- c(50, 100, 25, 75)

for (i in seq_along(transect_lengths)) {
  print(paste("Transect number", i, "measures", transect_lengths[i], "metres."))
}

##Tidyverse

#Using a standard for loop 
site_areas <- c(144, 400, 625)
results <- numeric(length(site_areas)) # Must build an empty container first

for(i in seq_along(site_areas)) {
  results[i] <- sqrt(site_areas[i])}
  
#Using the purr equivalent 
library(purrr)
mapped_results <- map_dbl(site_areas, sqrt) # Map directly, define expected output type explicitly

# Iterate a summary function over split lists of data
iris %>% 
  split(.$Species) %>% 
  map(~summary(.x))

```


```{r}
#I have a list of three vectors containing fish count data. Write a for-loop to calculate the mean of each, and then show me the exact parallel way to do it using purrr's map_dbl function.

fish_counts <- list(c(10, 15, 20), c(5, 8, 12), c(25, 30, 35))

# For-loop approach
means <- numeric(length(fish_counts))
for (i in seq_along(fish_counts)) 
  {
  means[i] <- mean(fish_counts[[i]])
  }

# Parallel approach with purrr
library(purrr)
mapped_means <- map_dbl(fish_counts, mean)

```

##FUNCTION CONSTRUCTION BLUEPRINT
```{r}

# Function Definition
calculate_coral_mortality <- function(initial_count, surviving_count) 
  {
  # Logic safety switch using our conditional tools!
  if (initial_count <= 0) 
    {
    stop("Initial count must be greater than zero.")
    }
  mortality_rate <- (initial_count - surviving_count) / initial_count
  return(mortality_rate)
  }

# Utilizing your custom function
calculate_coral_mortality(initial_count = 120, surviving_count = 84)
{
  if (initial_count <=0) {stop("Initial coun must be greater than zero.")}

mortality_rate <- (initial_count - surviving_count) / initial_count
return (mortalitry_rate) }


```


Build a custom function named convert_temp_c_to_f that accepts a single argument representing temperature in Celsius. The function should perform the math conversion: F = (C * 9/5) + 32. Include a conditional safety check using an if statement that halts execution via stop() if the entered value is below absolute zero (-273.15 degrees Celsius). Test your function with both a valid temperature and an invalid value to ensure the error handling behaves as expected.
```{r}
#build a function names convert_temp_c_to_f that accepts a single argument (temperature in Celsius) and performs F = (C* 9/5)+32. 



  

```


